\textbf{Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata}

### 2023-04-28 manuscript_v4/02-methods.qmd
need to fix figure to exponentiated

We used a logistic regression model to examine the impact of the Health Regions in vaccination rates while considering the socio-economic factors and and months covered by the survey (@tbl-descriptive-stats) and certain interactions (Race and Health Region and Race and income), as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a]. Because we identified differences in representativity between the survey data and the estimates from the Census, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using data from the Census and Health Region population totals;and fitted the regression model to the uncorrected and corrected data. Details regarding the correction can be found in the Appendix. All analyses were conducted in R 4.2.2 using the packages `survey` [@lumley2011],`tidyverse` [@wickham2019], `quarto` [@quarto], `modelsummary` [@modelsummary], and `gtsummary` [@gtsummary].

>>>>

Comparisons between the variables of the survey were performed using the $\chi^2$ test. We used a logistic regression model to examine the impact of the Health Regions in vaccination rates while considering the socio-economic factors and and months covered by the survey (@tbl-descriptive-stats) and certain interactions (Race and Health Region and Race and income), as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a]. Because we identified differences in representativity between the survey data and the estimates from the Census, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using data from the Census and Health Region population totals; and fitted the regression model to the uncorrected and corrected data. The reference groups in the model were: 16-34 years (age group), October(Month), CAD 60,000 and above (income), White/Caucasian (Race), Toronto (Health Region). Details regarding the correction can be found in the Appendix. All analyses were conducted in R 4.2.2 using the packages `survey` [@lumley2011],`tidyverse` [@wickham2019], `quarto` [@quarto], `modelsummary` [@modelsummary], and `gtsummary` [@gtsummary].

### 2023-04-28 manuscript_v4/03-results.qmd
need to fix figure to exponentiated

## Survey Results

@tbl-descriptive-stats shows the descriptive statistics from the Fields COVID-19 survey data for vaccination status and each of the covariates analyzed. The total number of entries analyzed was 3,549. Overall, 27\% (958) of survey respondents  reported not having received the first dose of the vaccine, whereas 73\% (2,591) reported having received it. Within each socio-economic factor, respondents who reported living in a household with an income under CAD 25,000 represented  37\%  (1307) of the number of entries, those within the CAD 25,000-59,999 income bracket represented 25% (889) of the total sample, and those with an income above CAD 60,000 represented 38 \% (1353) of the sample. Between the three income brackets, the lower vaccination rate corresponded to the under CAD 25,000 group, with 69\% .

Within the age groups of survey respondents, the age group between 16-34 years had the highest representation in the survey responses (1,520 or 42.8\% of all responses). Within this age bracket, 73\% of respondents indicated having received the vaccine, whereas the lowest vaccination rate was in the bracket of those 55 years of age and above, with 72\%. The Health Region with highest representation in the survey was Toronto, accounting for 1,323 entries (37.3\% of the total), with a vaccination rate of 72\%. Across the three months covered by the survey, the highest number of entries was in October of 2021, with a total of 1,732 respondents, which corresponded to 49\% of all the survey entries. Additionally, the proportion of respondents who reported being vaccinated was similar in all months, ranging between 72\% and 75\%, with the highest value corresponding to December of 2021.

Regarding race/ethnicity, individuals that identified as White/Caucasian represented 37\% of all entries (1,312) and had the highest vaccination uptake with 82\% of them indicating to have received the COVID-19 vaccine. On the other hand, the ethnic group with the lowest number of entries in the survey was Latin American, with a total of 180, or 5\% of all entries. Vaccination rates across all minority groups were below the value reported by White/Caucasians, with the lowest vaccination rate (60\%) being reported by individuals that identified as Indigenous.

@@

data_summary<-clean_data %>%
  tbl_summary(percent="row",
              by= first_dose,
              include=c(income_ord,
                        Age_group_ord,
                        Health_Region,
                        Month,
                        Race),
              label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group",
                 income_ord="Income (CAD)")
  ) %>% add_p()%>%
  bold_labels()

@@

@@

@tbl-model shows the results of the logistic regression models (for the uncorrected and corrected data) on vaccination status using socio-economic factors (age group, income, race), months covered by the survey (October, November, and December), geographical areas (Health Regions) and the interactions between income and race and Health Region and race. The corrected results show that there were no statistically significant differences in vaccination rate between the age groups or across the months covered by the survey as in both cases the odd ratios were similar. However, between the different household income brackets, individuals with an income under CAD 25,000 or between CAD 25,000-59,999 had significantly lower odds of vaccination than those with an income above CAD 60,000 (ORs=0.37 and 0.59, _p_<0.001 and _p_=0.010, respectively). Within Race/Ethnicity, individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.32, 0.32, 0.27, and _p_=0.003, _p_<0.001 and _p_=0.004, respectively); additionally, those individuals that reported to belong to the "Other" Race/Ethnicity group (which included the Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere groups according to the Census) had even lower odds of vaccination than the other minority groups (OR=0.22, _p_<0.001). Regarding Health Regions, individuals that reported living in the West Health Region (which comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent) had significantly higher odds of vaccination than those in the Health Region of Toronto (OR=1.54, _p_=0.031).

Moreover, statistically-significant odd ratios were determined in the case of the interaction of income and race; specifically, for individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.08, _p_=0.013), Black (OR=3.15, _p_=0.004), Latin American (OR=2.81, _p_=0.041), or that belonged to other minority groups  (OR=4.63, _p_<0.001). Within the CAD 25,000-59,999 income bracket, individuals who identified as belonging to other racial minority groups had significantly higher odds of vaccination (OR=6.96, _p_<0.001).

@@

```{r,model-uncorr-fig}
#| message: false
#| warning: false
#| echo: false
#| fig-align: left
#| fig-width: 7
#| fig-height: 7
#| label: fig-model-uncorr
#| fig-cap: "Coefficient estimates and confidence intervals for the uncorrected model. For the interaction terms, only those with a statistically significant values are shown."

@@

mplot1<-modelplot(m0,draw=FALSE,coef_omit = "Intercept") %>%
  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))

@@

```{r,model-corr-fig}
#| message: false
#| warning: false
#| echo: false
#| fig-align: left
#| fig-width: 7
#| fig-height: 7
#| label: fig-model-corr
#| fig-cap: "Coefficient estimates and confidence intervals for the corrected model. For the interaction terms, only those with a statistically significant values are shown."

@@

mplot2<-modelplot(m2,draw=FALSE,coef_omit = "Intercept") %>%
  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))

@@

mplot2 %>% 
ggplot(aes(y=term))+
  geom_vline(xintercept = 0, linetype="dashed",color="red") +
  geom_point(aes(x=estimate))+
  geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
  scale_y_discrete(limits=rev)+
  xlab("Coefficient estimates and 95% confidence intervals")+
  ylab("")+
  facet_wrap(~factor,ncol=1,strip.position = "right",scales = "free_y")+
  theme_classic()
```

>>>>

## Sample Characteristics

@tbl-descriptive-stats shows the characteristics of the data from the Fields COVID-19 survey used for analysis with regard to vaccination status. A total of 3,549 observations were used for analysis, from which 27\% (958) corresponded to individuals that reported not having received the first dose of the vaccine. Respondents between 16 and 34 years of age, and those that lived within the Health Region of Toronto had the highest number of observations in the data (42.8\% and 37.3\% of the total, respectively). Moreover, individuals who identified as White or Caucasian reported the highest vaccination uptake (82\%), in contrast to respondents who identified as which had had the lowest vaccination uptake rate (60\%) across all groups. From all the covariates, significant associations were identified between income, race, and vaccination rate.

@@

data_summary<-clean_data %>%
  tbl_summary(percent="row",
              by= first_dose,
              include=c(income_ord,
                        Age_group_ord,
                        Health_Region,
                        Month,
                        Race),
              label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group",
                 income_ord="Income (CAD)")
  ) %>% add_p(test = everything () ~ "chisq.test")%>%
  bold_labels()

@@

# test<-apply(xtabs(~first_dose_m + Race,clean_data), 2, chisq.test)
# 
# 
# clean_data %>%
#   filter(income_ord %in% c("60000 and above"))%>%
#   tbl_summary(percent="row",
#               by= first_dose_m,
#               missing="no",
#               include=Health_Region,
#               label=list(
#                  income_ord="Income (CAD)")
#   ) %>% add_p()%>%
#   bold_labels()

# clean_data %>%
#   filter(income_ord %in% c("25000-59999"))%>%
#   tbl_summary(percent="row",
#               by= income_ord,
#               missing="no",
#               include=(income_ord),
#               label=list(
#                  income_ord="Income (CAD)")
#   ) %>% add_p()%>%
#   bold_labels()

@@

@fig-model-uncorr shows the estimates from the logistic regression model of vaccination status for the uncorrected data, whereas the estimated obtained from the corrected data appear in @fig-model-corr. The results show significantly higher odds of vaccination in individuals with a high household income, and those that identified as White or Caucasian when compared with individuals in lower income brackets or that identified as part of equity-deserving groups.

Specifically, significantly lower odds of vaccination were identified for those with a household income under CAD 25,000 (OR=0.374, CI=[0.251,0.559]) and those with an income between CAD 25,000 and 59,999 (OR=0.586, CI=[0.390,0.882]).  Additionally, individuals who identified as Arab/Middle Eastern, Black, or Latin American, or that belonged to "Other" racial groups, which included the Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere groups according to the Census, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.31, 0.32, 0.27, 0.22, and CIs=[0.14,0.68], [0.17,0.60], [0.11,0.66], and [0.12,0.41], respectively). Regarding Health Regions, individuals that reported living in the West Health Region (which includes the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent) had significantly higher odds of vaccination than those in the Health Region of Toronto (OR=1.54, CI=[1.04,2.29]).

Interestingly, higher odds of vaccination were also estimated in the case of individuals in equity-deserving groups with a household income below CAD 25,000 when compared to those with a household income above CAD 60,000.  This held true in the case of Arab/Middle Eastern (OR=3.08, _p_=0.013), Black (OR=3.15, _p_=0.004), and Latin American (OR=2.81, _p_=0.041) individuals, as well as respondents who belonged to "Other" minority groups  (OR=4.63, _p_<0.001). Within the CAD 25,000-59,999 income bracket, individuals who identified as belonging to other racial minority groups had significantly higher odds of vaccination (OR=6.96, _p_<0.001).

@@

```{r,model-uncorr-fig}
#| message: false
#| warning: false
#| echo: false
#| fig-align: left
#| fig-width: 7
#| fig-height: 7
#| label: fig-model-uncorr
#| fig-cap: "Coefficient estimates and confidence intervals for the uncorrected model. Only statistically significant interaction terms are shown. Full estimates can be found in the Appendix."

@@

mplot1<-modelplot(m0,draw=FALSE,
                  coef_omit = "Intercept", exponentiate = TRUE) %>%
  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))

@@

```{r,model-corr-fig}
#| message: false
#| warning: false
#| echo: false
#| fig-align: left
#| fig-width: 7
#| fig-height: 7
#| label: fig-model-corr
#| fig-cap: "Coefficient estimates and confidence intervals for the corrected model. Only statistically significant interaction terms are shown. Full estimates can be found in the Appendix."

@@

mplot2<-modelplot(m2,draw=FALSE,coef_omit = "Intercept",exponentiate = TRUE) %>%
  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))

@@

mplot2 %>% 
ggplot(aes(y=term))+
  geom_vline(xintercept = 5.73, linetype="dashed",color="red") +
  geom_point(aes(x=estimate))+
  geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
  scale_y_discrete(limits=rev)+
  xlab("Coefficient estimates and 95% confidence intervals")+
  ylab("")+
  facet_wrap(~factor,ncol=1,strip.position = "right",scales = "free_y")+
  theme_classic()
```

### 2023-04-28 manuscript_v4/main.tex
need to fix figure to exponentiated

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

We used a logistic regression model to examine the impact of the Health
Regions in vaccination rates while considering the socio-economic
factors and and months covered by the survey
(Table~\ref{tbl-descriptive-stats}) and certain interactions (Race and
Health Region and Race and income), as previous studies have shown that
socio-economic factors and their interactions are significant predictors
of intent of vaccination and vaccination
status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{32}--\protect\hyperlink{ref-cnat2022a}{34}}.
Because we identified differences in representativity between the survey
data and the estimates from the Census, we used an iterative
proportional fitting procedure
(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{35}}
to correct the data using data from the Census and Health Region
population totals;and fitted the regression model to the uncorrected and
corrected data. Details regarding the correction can be found in the
Appendix. All analyses were conducted in R 4.2.2 using the packages
\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{36}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{37}},
\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{38}},
\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{39}},
and
\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{40}}.

@@

\hypertarget{survey-results}{%
\subsection{Survey Results}\label{survey-results}}

Table~\ref{tbl-descriptive-stats} shows the descriptive statistics from
the Fields COVID-19 survey data for vaccination status and each of the
covariates analyzed. The total number of entries analyzed was 3,549.
Overall, 27\% (958) of survey respondents reported not having received
the first dose of the vaccine, whereas 73\% (2,591) reported having
received it. Within each socio-economic factor, respondents who reported
living in a household with an income under CAD 25,000 represented 37\%
(1307) of the number of entries, those within the CAD 25,000-59,999
income bracket represented 25\% (889) of the total sample, and those
with an income above CAD 60,000 represented 38 \% (1353) of the sample.
Between the three income brackets, the lower vaccination rate
corresponded to the under CAD 25,000 group, with 69\% .

Within the age groups of survey respondents, the age group between 16-34
years had the highest representation in the survey responses (1,520 or
42.8\% of all responses). Within this age bracket, 73\% of respondents
indicated having received the vaccine, whereas the lowest vaccination
rate was in the bracket of those 55 years of age and above, with 72\%.
The Health Region with highest representation in the survey was Toronto,
accounting for 1,323 entries (37.3\% of the total), with a vaccination
rate of 72\%. Across the three months covered by the survey, the highest
number of entries was in October of 2021, with a total of 1,732
respondents, which corresponded to 49\% of all the survey entries.
Additionally, the proportion of respondents who reported being
vaccinated was similar in all months, ranging between 72\% and 75\%,
with the highest value corresponding to December of 2021.

Regarding race/ethnicity, individuals that identified as White/Caucasian
represented 37\% of all entries (1,312) and had the highest vaccination
uptake with 82\% of them indicating to have received the COVID-19
vaccine. On the other hand, the ethnic group with the lowest number of
entries in the survey was Latin American, with a total of 180, or 5\% of
all entries. Vaccination rates across all minority groups were below the
value reported by White/Caucasians, with the lowest vaccination rate
(60\%) being reported by individuals that identified as Indigenous.

@@

\textbf{?@tbl-model} shows the results of the logistic regression models
(for the uncorrected and corrected data) on vaccination status using
socio-economic factors (age group, income, race), months covered by the
survey (October, November, and December), geographical areas (Health
Regions) and the interactions between income and race and Health Region
and race. The corrected results show that there were no statistically
significant differences in vaccination rate between the age groups or
across the months covered by the survey as in both cases the odd ratios
were similar. However, between the different household income brackets,
individuals with an income under CAD 25,000 or between CAD 25,000-59,999
had significantly lower odds of vaccination than those with an income
above CAD 60,000 (ORs=0.37 and 0.59, \emph{p}\textless0.001 and
\emph{p}=0.010, respectively). Within Race/Ethnicity, individuals who
identified as Arab/Middle Eastern, Black, or Latin American, had
significantly lower odds of vaccination than those in the
White/Caucasian group (ORs=0.32, 0.32, 0.27, and \emph{p}=0.003,
\emph{p}\textless0.001 and \emph{p}=0.004, respectively); additionally,
those individuals that reported to belong to the ``Other''
Race/Ethnicity group (which included the Southeast Asian, Filipino, West
Asian, and Minorities Not Identified Elsewhere groups according to the
Census) had even lower odds of vaccination than the other minority
groups (OR=0.22, \emph{p}\textless0.001). Regarding Health Regions,
individuals that reported living in the West Health Region (which
comprises the regions of Waterloo and Niagara, the counties of
Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand,
Brant, and Chatham-Kent) had significantly higher odds of vaccination
than those in the Health Region of Toronto (OR=1.54, \emph{p}=0.031).

@@

\caption{\label{fig-model-uncorr}Coefficient estimates and confidence
intervals for the uncorrected model. For the interaction terms, only
those with a statistically significant values are shown.}

@@

\caption{\label{fig-model-corr}Coefficient estimates and confidence
intervals for the corrected model. For the interaction terms, only those
with a statistically significant values are shown.}

>>>>

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, sharp corners, interior hidden]}{\end{tcolorbox}}\fi

@@

Comparisons between the variables of the survey were performed using the
\(\chi^2\) test. We used a logistic regression model to examine the
impact of the Health Regions in vaccination rates while considering the
socio-economic factors and and months covered by the survey
(Table~\ref{tbl-descriptive-stats}) and certain interactions (Race and
Health Region and Race and income), as previous studies have shown that
socio-economic factors and their interactions are significant predictors
of intent of vaccination and vaccination
status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{32}--\protect\hyperlink{ref-cnat2022a}{34}}.
Because we identified differences in representativity between the survey
data and the estimates from the Census, we used an iterative
proportional fitting procedure
(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{35}}
to correct the data using data from the Census and Health Region
population totals; and fitted the regression model to the uncorrected
and corrected data. The reference groups in the model were: 16-34 years
(age group), October(Month), CAD 60,000 and above (income),
White/Caucasian (Race), Toronto (Health Region). Details regarding the
correction can be found in the Appendix. All analyses were conducted in
R 4.2.2 using the packages
\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{36}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{37}},
\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{38}},
\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{39}},
and
\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{40}}.

@@

\hypertarget{sample-characteristics}{%
\subsection{Sample Characteristics}\label{sample-characteristics}}

Table~\ref{tbl-descriptive-stats} shows the characteristics of the data
from the Fields COVID-19 survey used for analysis with regard to
vaccination status. A total of 3,549 observations were used for
analysis, from which 27\% (958) corresponded to individuals that
reported not having received the first dose of the vaccine. Respondents
between 16 and 34 years of age, and those that lived within the Health
Region of Toronto had the highest number of observations in the data
(42.8\% and 37.3\% of the total, respectively). Moreover, individuals
who identified as White or Caucasian reported the highest vaccination
uptake (82\%), in contrast to respondents who identified as which had
had the lowest vaccination uptake rate (60\%) across all groups. From
all the covariates, significant associations were identified between
income, race, and vaccination rate.

@@

Figure~\ref{fig-model-uncorr} shows the estimates from the logistic
regression model of vaccination status for the uncorrected data, whereas
the estimated obtained from the corrected data appear in
Figure~\ref{fig-model-corr}. The results show significantly higher odds
of vaccination in individuals with a high household income, and those
that identified as White or Caucasian when compared with individuals in
lower income brackets or that identified as part of equity-deserving
groups.

Specifically, significantly lower odds of vaccination were identified
for those with a household income under CAD 25,000 (OR=0.374,
CI={[}0.251,0.559{]}) and those with an income between CAD 25,000 and
59,999 (OR=0.586, CI={[}0.390,0.882{]}). Additionally, individuals who
identified as Arab/Middle Eastern, Black, or Latin American, had
significantly lower odds of vaccination than those in the
White/Caucasian group (ORs=0.32, 0.32, 0.27, and \emph{p}=0.003,
\emph{p}\textless0.001 and \emph{p}=0.004, respectively); additionally,
those individuals that reported to belong to the ``Other''
Race/Ethnicity group (which included the Southeast Asian, Filipino, West
Asian, and Minorities Not Identified Elsewhere groups according to the
Census) had even lower odds of vaccination than the other minority
groups (OR=0.22, \emph{p}\textless0.001). Regarding Health Regions,
individuals that reported living in the West Health Region (which
comprises the regions of Waterloo and Niagara, the counties of
Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand,
Brant, and Chatham-Kent) had significantly higher odds of vaccination
than those in the Health Region of Toronto (OR=1.54, \emph{p}=0.031).

@@

\caption{\label{fig-model-uncorr}Coefficient estimates and confidence
intervals for the uncorrected model. Only statistically significant
interaction terms are shown. Full estimates can be found in the
Appendix.}

@@

\caption{\label{fig-model-corr}Coefficient estimates and confidence
intervals for the corrected model. Only statistically significant
interaction terms are shown. Full estimates can be found in the
Appendix.}
